%---------------------------------------- bibliography
\begin{thebibliography}{99}
\bibitem{thesis} Beaugendre, H.,  \textit{A PDE-Based 3D Approach to In-Flight Ice Accretion}, McGill University. Montreal, Québec. Junio 2003

\bibitem{beta} Bilodeau, David. \textit{An Eulerian Framework for Supercooled Large Droplets Splashing and Bouncing.} DOI: 10.13140/RG.2.1.2768.2327. 2014.

\bibitem{Pena} Pena, D., Hoarau, Y. \& Laurendeau, E., \textit{A single step ice accretion model using Level-Set method.} Journal of Fluids and Structures, 65, pp.278-294. 2016.

\bibitem{messinger} Messinger, B.,  \textit{Equilibrium Temperature of an Unheated Icing Surface as a Function of Air Speed.} Journal of the Aeronautical Sciences, 20(1), pp.29-42. 1953.

\bibitem{hedde} Hedde, T., Guffond, D.,\textit{ ONERA Three-Dimensional Icing Model}, AIAA Journal, Vol. 33, Nº. 6. Junio 1995

\bibitem{toro} Toro, E.,  \textit{Riemann Solvers and Numerical Methods for Fluid Dynamic}, Springer. Berlin, 2009.

\bibitem{ruff} Ruff, G. A., \& Berkowitz, M., \textit{Users Manual for the NASA Lewis Ice Accretion Prediction Code (LEWICE)}, NASA Contractor Report 185129, Mayo 1990

\bibitem{Heat_Transfer} Kreith, F., Manglik, R.\& Bohn, M.,  \textit{Principles Of Heat Transfer. }7th ed., Cengage Learning. Stamford, 2017. 

\bibitem{JAHR} Hernández Ramos, J.,  \textit{Cálculo Numérico En Ecuaciones Diferenciales Ordinarias.} 2nd ed, Independiente. Madrid, 2018.

\bibitem{jahr_eusebio} Hernández Ramos, J. \& Valero Sánchez, E.,  \textit{Análisis Y Cálculo Numérico En Ecuaciones En Derivadas Parciales.}  Escuela Técnica Superior de Ingenieros Aeronáuticos. Madrid, 2002.

\bibitem{libro_termo} de Assas Martínez de Morentín, P., Barrero Gil, A.\& Pérez Grande, I., \textit{Termodinámica: Teoría.} , Sección de Publicaciones, ETSIAE. Madrid, 2017.

\bibitem{hirsch} Hirsch, C.,  \textit{Numerical Computation Of Internal And External Flows.} 2nd ed., Wiley. Chichester, 2002. 

\bibitem{politovich} Politovich, M., \textit{Aircraft Icing.} Encyclopedia of Atmospheric Sciences, pp.68-75. 2003. 

\bibitem{Icing_effects} Pouryoussefi, S., Mirzaei, M., Nazemi, M., Fouladi, M.\& Doostmahmoudi, A.,  \textit{Experimental study of ice accretion effects on aerodynamic performance of an NACA 23012 airfoil}. Chinese Journal of Aeronautics, 29(3), pp.585-595. 2016.

\bibitem{mesoscale} Tremblay, A., Cober, S., Glazer, A., Isaac, G.\& Mailhot, J., 1996. \textit{An Intercomparison of Mesoscale Forecasts of Aircraft Icing Using SSM/I Retrievals.} Weather and Forecasting, 11(1), pp.66-77.
\bibitem{meteorology} Potter, T.\& Colman, B.,  \textit{Handbook Of Weather, Climate And Water.}  Wiley. New York, 2003.

\bibitem{avion_y_piloto} Viñas, J.,  \textit{El engelamiento.} Avión \& Piloto, 18, pp.32-35. 2013.

\bibitem{enemigo_oculto} Flores Gustavo, A. \textit{Engelamiento, un enemigo oculto entre las nubes}, J.I.A.A.C, Departamento Meteorología Aeronáutica de la DGSMN, 2011.

\bibitem{tema_14} Toscano Jiménez, M. \textit{Tema 14: El engelamiento} [Diapositivas de PowerPoint]. Recuperado 25 de mayo 2020, de \url{http://www.esi2.us.es/DFA/Meteo/TEMA\%2014,\%20v130515.pdf}

\bibitem{TAU_code} Schwamborn, D., Gerhold, T. \& Heinrich, R. \textit{The DLR TAU-code: Recent applications in research and industry.}, DLR, ECCOMAS CFD 2006 CONFERENCE

\bibitem{TAU_figura} Schütte, A., Boelens, O., Loeser, T. \& Oehlke, M.,  \textit{Prediction of the Flow around the X-31 Aircraft using two different CFD methods.} 28th AIAA Applied Aerodynamics Conference. 2010.

%%-------------------PÁGINAS WEB--------------------
\bibitem{FAR_icing} United States. Federal Aviation Administration. \textit{Federal Aviation Regulations.} Part 25, Appendix  C, Airworthiness Standards: Transport Category Planes. Atmospheric Icing Conditions. Washington, D.C. :U.S. Dept. of Transportation, Federal Aviation Administration : , 1989. 

\bibitem{Icing_accidents} Eick, D.,\textit{Aircraft Icing Accidents}. [online] Ral.ucar.edu. Disponible en: \url{https://ral.ucar.edu/sites/default/files/public/events/2015/in-flight-icing-users-technical-interchange-meeting-tim/docs/eick-ntsb-ncar-icing-presentation.pdf} .

\bibitem{compressor_icing} Banke, J. \textit{Airplane Plus Heat Plus Ice Equals Mystery.} [online] NASA. Disponible en: \url{https://www.nasa.gov/topics/aeronautics/features/ice_mystery.html}. 2017

\bibitem{Tailplane} Skybrary.aero. \textit{Ice Contaminated Tailplane Stall} [online] Disponible en: \url{https://www.skybrary.aero/index.php/Ice_Contaminated_Tailplane_Stall}.

\bibitem{Freezing_rain} Skybrary.aero. \textit{Freezing Rain}[online] Disponible en: \url{https://www.skybrary.aero/index.php/Freezing_Rain}.

\bibitem{inlet} Cleynen, O., \textit{Icing On The Inlet Of A CFM56 Turbofan Engine} [online] es.m.wikipedia.org. Disponible en: \url{https://es.m.wikipedia.org/wiki/Archivo:Icing_on_the_inlet_of_a_CFM56_turbofan_engine_(2).jpg}.

\bibitem{carburetor} Faasafety.gov. Learning Center Courses Content - FAA - Faasteam - Faasafety.Gov. [online] Disponible en: \url{https://www.faasafety.gov/gslac/ALC/course_content_popup.aspx?preview=true&cID=33&sID=150}.

\bibitem{cherry} Tikkanen, A., \textit{Glaze | Meteorology}. [online] Encyclopedia Britannica. Disponible en: \url{https://www.britannica.com/science/glaze}.

\bibitem{icing_glenn} Glenn Research Center \textit{| NASA. 2020. Icing Research | Glenn Research Center | NASA.} [online] Disponible en: \url{https://www1.grc.nasa.gov/aeronautics/icing/\# gallery}

\bibitem{aguanieve} Cherneski, D., Chapman, R., Robe, F. Lundgren, J.  \textit{Forecasting System for Snow \& Ice Accretion on Cable Stay Bridges and Power Lines}. World Weather Open Science Conference 2014
Session: SCI-PS152.04.

\bibitem{Vibrating_rod} Intelligentvisionsys.com.  \textit{Current Aircraft Icing Detection | Intelligent Vision Systems}. [online] Disponible en: \url{https://www.intelligentvisionsys.com/technology/a-sampling-of-current-aircraft-icing-detection-technologies/} 

\bibitem{Capacitance} Zhi, X., Cho, H., Wang, B., Ahn, C., Moon, H.\& Go, J.,  \textit{Development of a Capacitive Ice Sensor to Measure Ice Growth in Real Time}. Sensors, 15(3), pp.6688-6698. 2015.

\bibitem{De_icing_airport} en.wikipedia.org.  \textit{De-Icing.} [online] Disponble en: \url{https://en.wikipedia.org/wiki/De-icing}. 

\bibitem{Pneumatic_boot} Aircraft Systems,  \textit{Aircraft Wing And Stabilizer Deicing Systems}. [online] Aircraftsystemstech.com. Disponible en: \url{https://www.aircraftsystemstech.com/2017/05/wing-and-stabilizer-deicing-systems.html}. 

\bibitem{Piccolo} English.iet.cas.cn.  \textit{Experimental Study Of Jet Impingement Heat Transfer On A Variable-Curvature Concave Surface In A Wing Leading Edge} [online] Disponible en: \url{http://english.iet.cas.cn/Research/ResearchProgress/201711/t20171121_186386.html}. 
\bibitem{nanotubos} Nanowerk.  \textit{Carbon Nanotubes In Coating Make A Heatable Paint To Prevent Ice Buildup.} [online] Disponible en: \url{https://www.nanowerk.com/news/newsid=14510.php}. 

\bibitem{nanotech} García Barco, V.,  \textit{FORMACIÓN DE HIELO Y NANOTECNOLOGÍA EN AERONÁUTICA} - Hispaviación. [online] Hispaviación. Disponible en: \url{http://www.hispaviacion.es/formacion-de-hielo-y-nanotecnologia-en-aeronautica/} .
\end{thebibliography}




